Sound and Automated Synthesis of Digital Stabilizing Controllers for Continuous Plants \thanks{Supported by EPSRC grant EP/J012564/1, ERC project 280053 (CPROVER) and the H2020 FET OPEN SC$^2$.}

### 2017-02-12 hscc2017/paper.tex
addressed some of Alessandro's comments

%------------------------------------
\subsection{Verifying Closed-loop Control Systems \\ \blue{Effects of Imprecisions, Quantization, and Finite Word Length Representation}}
\label{verifying-closed-loop-control-systems}
%------------------------------------

@@

CEGIS-based control synthesis requires a formal verifier to check whether a
candidate controller meets the requirements when combined with the plant. 
We use the Digital-System Verifier (DSVerifier)~\cite{IsmailBCFF15} \red{[unclear difference btw dssynth and dsverifier]} for this
purpose.  It checks the stability of closed-loop control systems and
considers finite-word length (FWL) effects in the digital controller, and
uncertainty parameters in the plant model (plant intervals)~\cite{Bessa16}. 
Let $C(z)$ be a digital controller and $G(z)$ be a digital representation of the plant, 
given as
%
\begin{align}
\small
\label{controller_plant_tf}
C(z)&=\frac{C_n(z)}{C_d(z)}=\frac{\beta_{0}+\beta_{1}z^{-1}+...+\beta_{M_C}z^{-M_C}}{\alpha_{0}+\alpha_{1}z^{-1}+...+\alpha_{N_C}z^{-N_C}}, \\
%\label{plant_tf}
G(z)&=\frac{G_n(z)}{G_d(z)}=\frac{b_{0}+b_{1}z^{-1}+...+b_{M_G}z^{-M_G}}{a_{0}+a_{1}z^{-1}+...+a_{N_G}z^{-N_G}}.
\end{align}
%
where $\vec{\beta}$ and $\vec{\alpha}$ are vectors containing the
controller's coefficients; similarly, $\vec{b}$ and $\vec{a}$ denote the
plant's coefficients; 
and finally $N_{(\cdot)}$ and $M_{(\cdot)}$ indicate the order of the polynomials,  
and we require in particular that $N_G \geq M_G$.

Uncertainties in $G(z)$ may appear due to: 
1. uncertainties in $G(s)$ (we denote them $\Delta_p{G}_n(s),\Delta_p{G}_d(s)$) arising from tolerances/imprecision in the original model; 
2. errors in the numerical calculations (\red{denoted \ldots}); 
and 3. roundoffs in the quantization (\red{denoted \ldots}).  
These uncertainties are parametrically expressed by additive terms, eventually resulting in an uncertain model $\hat{G}(z)$, such that:
%
\begin{equation}
\hat{G}(z)=\frac{G_n(z)+\Delta G_n(z)}{G_d(z)+\Delta G_d(z)},
\end{equation}
%
%\red{[what is the relationship between the deltas in this formula, and those in the previous paragraph? please clarify and hopefully simplify formulae. ]}
which will be represented by the following transfer function:
%
\begin{equation}
\hat{G}(z)=\frac{\hat{b}_{0}+\hat{b}_{1}z^{-1}+...+\hat{b}_{M_G}z^{-M_G}}{\hat{a}_{0}+\hat{a}_{1}z^{-1}+...+\hat{a}_{N_G}z^{-N_G}}. 
\end{equation}
%
%\red{[why is this condition on the order of the polynomials raised only now? how is it used later?]}
Notice that, due to the nature of the methods we use for the stability check, we require
that the parametric errors in the plant have the same polynomial order as
the plant itself (\blue{indeed, all other errors described in this paper are ensured to
fulfil this property in our implementation}). 
%\red{[have we later argued that this is indeed the case?]}) 
%\blue{and that enter the plant linearly}.

@@

%\red{[why is the output in Figure 2 tilde, whereas that in  figure 1 hat?]}
%
In Figure~\ref{fig:digsystem1}, two sources of uncertainty are illustrated:
parametric uncertainties related to modeling errors  
(which are represented by~$\Delta_p \vec{G}$),  
and uncertainties introduced by quantizations in the ADC and
DAC conversions \red{[how are these denoted?]}.  
These sources are assumed to be non-deterministic \blue{[here you mean specifically the quantization effects?]}:  
%which poses no further restriction to our model, 
thus, 
the uncertain model may be rewritten as a vector of coefficients in the z-domain using equation (\ref{eq:hatgvector}) as 
%
\begin{equation}
\label{eq:complete_unc_model}
\vec{\hat{G}}=\vec{G}+\Delta_p \vec{G}. 
\end{equation}
%
%where $\Delta_p \vec{G}$ is the term that corresponds to the parametric uncertainties. 
Recall that we discussed how the quantization noise is encompassed as an additive term, 
which means it does not enter parametrically in the transfer function. 
Instead, we later show that the system is stable given these non-deterministic noise inputs.

@@

Let $\hat{C}(z)$ be the digital controller transfer function represented
using this FWL with integer size $I$ and fractional size~$F$; $I$ affects
the range of the representation and is set to avoid overflows, while $F$
affects the precision and the truncation after arithmetic operations. 
\red{[we ought to align the following notations with those use in the CAV submission.]}
We shall denote the FWL domain of the coefficients by $\mathbb{R}\langle I,F
\rangle$
%
%Since the controller is implemented in the $\mathbb{R}\langle I,F \rangle$ 
%domain, {\it i.e.} a Fixed-Point Finite Word Length with integer size $I$ and fractional
%size $F$, we shall denote this domain as $\mathbb{R}\langle I,F \rangle$ 
%
and assume that there exists a function
$\mathcal{F}_n\langle I,F \rangle(\cdot):\mathcal{P}^{n}\rightarrow \mathcal{P}^{n}\langle I,F \rangle$,
which applies the FWL effects to a polynomial, with an equivalent $\mathcal{F}_{n,m}\langle I,F \rangle(\cdot):\mathcal{P}^{n,m}\rightarrow \mathcal{P}^{n,m}\langle I,F \rangle$ that applies the same effect to a ratio of polynomials, where
$\mathcal{P}^{n}$ is the
space of polynomials of $n$-th order, $\mathcal{P}^{n}\langle I,F
\rangle$ is the space of polynomials with coefficients in
$\mathbb{R}\langle I,F \rangle$, and $\mathcal{P}^{n,m}$,  $\mathcal{P}^{n,m}\langle I,F
\rangle$ are their rational equivalent.

@@

%------------------------------------
\subsection{Stability under FWL Representation and under Quantisation Noise \blue{(and params uncertainties?)}}
\label{sec:stability}
%------------------------------------

Sound synthesis of the digital controller requires the consideration of the effect of FWL on the controller, 
and of quantization noises in the closed-loop system.  
Let the quantizer $Q1$ \blue{[new notation- needed really?] (DAC)} be the source of a white noise $\nu_{1}$, 
and $Q2$ \blue{[ditto] (ADC)} be the source of a white noise $\nu_{2}$.  The following equation
models the system in Figure~\ref{fig:sampledsystem}, encompassing the
parametric uncertainties $\Delta \vec{G}$ and the FWL effects on
the controller $\hat{C}(z)$:
%
\begin{equation}
\hat{Y}(z)=\nu_{1}(z)+\hat{G}(z)\cdot C(z)\cdot R(z)+\hat{G}(z)\cdot\nu_{2}(z)-\hat{G}(z)\cdot \hat{C}(z)\cdot \hat{Y}(z). 
\end{equation}
%
The above can be rewritten as follows:
%
\begin{equation}
\label{eq:outputfunctions}
\hat{Y}(z)=H_{1}(z)\cdot\nu_{1}(z)+H_{2}(z)\cdot\nu_{2}(z)+H_{3}(z)\cdot R(z),
\end{equation}
%
where
%
\begin{equation*}
H_{1}(z)=\frac{1}{1+\hat{G}(z)\cdot \hat{C}(z)}, 
\end{equation*}
%
\begin{equation*}
H_{2}(z)=\frac{\hat{G}(z)}{1+\hat{G}(z)\cdot \hat{C}(z)}, 
\end{equation*}
%
\begin{equation*}
H_{3}(z)=\frac{\hat{G}(z)\cdot \hat{C}(z)}{1+\hat{G}(z)\cdot \hat{C}(z)}. 
\end{equation*}

\begin{myassumption}
\label{whitenoise}
%
The quantization noises $\nu_{1}$ \blue{(from $Q1$)} and $\nu_{2}$ \blue{(from $Q2$)} are
uncorrelated white noises and their amplitudes are always bounded by the
half of quantization step~\cite{astrom1997computer}, i.e., $\vert \nu_{1}
\vert \leq \frac{q_{1}}{2}$ and $\vert \nu_{2} \vert \leq \frac{q_{2}}{2}$,
where $q_{1}$ and $q_{2}$ are the quantization steps of ADC and DAC, respectively.
% 
\end{myassumption}

@@

Next, we describe the specific property that we pass to the program
synthesizer as the specification $\sigma$.  There are two different
algorithms in \red{DSVerifier} that can be used for stability
verification~\cite{daes20161, Bessa16}, one based on Schur's decomposition
and another one based on Jury's criterion~\cite{astrom1997computer}.  Here,
we choose Jury's method~\cite{astrom1997computer}, due to its efficiency,
and we use this method to check the stability in the $z$-domain for the
characteristic polynomial $S(z)$ defined in~\eqref{eq:internal_stab_lemma}.
%
% of the matrix $\left( \begin{array}{c} Y \\ U \end{array}\right)$, where $Y(z)$ is the closed-loop system output and U(z) is the controller output. 
%% \red{[unclear: what matrix? We need to clarify the relationship
%% between the char. poly. $S(z)$ and the later quantities $\Delta
%% N_{G}(z), \Delta D_{G}(z)$ and corresponding controller's
%% structure.]}.
%
We~consider the following form for $S(z)$:
%
\begin{equation*}
S(z) = a_0z^N+a_1z^{N-1}+\cdots+a_{N-1}z+a_N=0, a_0\neq0. 
\end{equation*}
%\cdsay{is this for the matrix $\left( \begin{array}{c} Y \\ U \end{array}\right)$?}

@@

The {\sc synthesize} phase uses BMC to 
compute a solution $C(z)$. % An important remark is that this 
% controller is synthesised in the $\mathbb{R}\langle I,F \rangle$
% domain.
There are two approaches for the {\sc verify} phase.
% based on the described model.  
The first approach uses interval arithmetic \cite{moore1966interval}
to represent the coefficients
$[{c}_i-\Delta_p{c}_i-\Delta_b{c}_i,\,\allowbreak
{c}_i+\Delta_p{c}_i-\Delta_b{c}_i+(2^{-F_p})]$ 
and rounds operations outwards during calculations.  This effectively allows
us to simultaneously evaluate the whole collection of plants $\mathfrak{G}$ \red{[weird new notation needed? used elsewhere?]}
plus the effects of numeric calculations in the engine.  Synthesized
controllers using this verifier will always be stable for all plants in the
family.
% and only generate counterexamples when the precision of the
%engine is insufficient (\emph{ie} no stable plant could be found). 
Our preliminary experiments showed that a synthesis approach
using this verification engine has poor performance and we
%Given the high computational cost of this
%approach, 
therefore designed a second approach, which we describe below.
Our experimental results described in Section~\ref{sec:experiments}
%try both approaches and provide the time for the fastest. 
show that the speedup yielded by the second approach 
is in most cases of at least two orders of magnitude.

>>>>

%------------------------------------
\subsection{Verifying Closed-loop Control Systems \\ \blue{Under the Effect of Imprecisions, Quantization, and Finite Word Length Representation}}
\label{verifying-closed-loop-control-systems}
%------------------------------------

@@

CEGIS-based control synthesis requires a formal verifier to check whether a
candidate controller meets the requirements when combined with the plant. 
We use the Digital-System Verifier (DSVerifier)~\cite{IsmailBCFF15} in the
verification module for \tool.  It checks the stability of closed-loop control systems and
considers finite-word length (FWL) effects in the digital controller, and
uncertainty parameters in the plant model (plant intervals)~\cite{Bessa16}. 
Let $C(z)$ be a digital controller and $G(z)$ be a digital representation of the plant, 
given as
%
\begin{align}
\small
\label{controller_plant_tf}
C(z)&=\frac{C_n(z)}{C_d(z)}=\frac{\beta_{0}+\beta_{1}z^{-1}+...+\beta_{M_C}z^{-M_C}}{\alpha_{0}+\alpha_{1}z^{-1}+...+\alpha_{N_C}z^{-N_C}}, \\
%\label{plant_tf}
G(z)&=\frac{G_n(z)}{G_d(z)}=\frac{b_{0}+b_{1}z^{-1}+...+b_{M_G}z^{-M_G}}{a_{0}+a_{1}z^{-1}+...+a_{N_G}z^{-N_G}}.
\end{align}
%
where $\vec{\beta}$ and $\vec{\alpha}$ are vectors containing the
controller's coefficients; similarly, $\vec{b}$ and $\vec{a}$ denote the
plant's coefficients; 
and finally $N_{(\cdot)}$ and $M_{(\cdot)}$ indicate the order of the polynomials,  
and we require in particular that $N_G \geq M_G$.

Uncertainties in $G(z)$ may appear due to: 
1. uncertainties in $G(s)$ (we denote them $\Delta_p{G}_n(s),\Delta_p{G}_d(s)$) arising from tolerances/imprecision in the original model; 
2. errors in the numerical calculations (equal to $0$ since we introduce them as external noises $\nu_1$ and $\nu_2$); 
and 3. roundoffs in the quantization (denoted $\Delta_b{G}_n(s),\Delta_b{G}_d(s)$).  
These uncertainties are parametrically expressed by additive terms, eventually resulting in an uncertain model $\hat{G}(z)$, such that:
%
\begin{equation}
\hat{G}(z)=\frac{G_n(z)+\Delta G_n(z)}{G_d(z)+\Delta G_d(z)},
\end{equation}
%
%\red{[what is the relationship between the deltas in this formula, and those in the previous paragraph? please clarify and hopefully simplify formulae. ]}
which will be represented by the following transfer function:
%
\begin{equation}
\hat{G}(z)=\frac{\hat{b}_{0}+\hat{b}_{1}z^{-1}+...+\hat{b}_{M_G}z^{-M_G}}{\hat{a}_{0}+\hat{a}_{1}z^{-1}+...+\hat{a}_{N_G}z^{-N_G}}. 
\end{equation}
%
%\red{[why is this condition on the order of the polynomials raised only now? how is it used later?]}
Notice that, due to the nature of the methods we use for the stability check, we require
that the parametric errors in the plant have the same polynomial order as
the plant itself (\blue{indeed, all other errors described in this paper are ensured to
fulfil this property in our implementation}). 
%\red{[have we later argued that this is indeed the case?]}) 
%\blue{and that enter the plant linearly}.

@@

%\red{[why is the output in Figure 2 tilde, whereas that in  figure 1 hat?]}
%
In Figure~\ref{fig:digsystem1}, two sources of uncertainty are illustrated:
parametric uncertainties related to modeling errors  
(which are represented by~$\Delta_p \vec{G}$),  
and uncertainties introduced by quantizations in the ADC and
DAC conversions ($\nu_1$ and $\nu_2$).  
These sources are assumed to be non-deterministic \blue{[here you mean specifically the quantization effects?]}:  
%which poses no further restriction to our model, 
thus, 
the uncertain model may be rewritten as a vector of coefficients in the z-domain using equation (\ref{eq:hatgvector}) as 
%
\begin{equation}
\label{eq:complete_unc_model}
\vec{\hat{G}}=\vec{G}+\Delta_p \vec{G}. 
\end{equation}
%
%where $\Delta_p \vec{G}$ is the term that corresponds to the parametric uncertainties. 
Recall that we discussed how the quantization noise is encompassed as an additive term, 
which means it does not enter parametrically in the transfer function. 
Instead, we later show that the system is stable given these non-deterministic noise inputs.

@@

Let $\hat{C}(z)$ be the digital controller transfer function represented
using this FWL with integer size $I$ and fractional size~$F$; $I$ affects
the range of the representation and is set to avoid overflows, while $F$
affects the precision and the truncation after arithmetic operations. 
We shall denote the FWL domain of the coefficients by $\mathbb{R}\langle I,F
\rangle$
%
%Since the controller is implemented in the $\mathbb{R}\langle I,F \rangle$ 
%domain, {\it i.e.} a Fixed-Point Finite Word Length with integer size $I$ and fractional
%size $F$, we shall denote this domain as $\mathbb{R}\langle I,F \rangle$ 
%
and assume that there exists a function
$\mathcal{F}_n\langle I,F \rangle(\cdot):\mathcal{P}^{n}\rightarrow \mathcal{P}^{n}\langle I,F \rangle$,
which applies the FWL effects to a polynomial, with an equivalent $\mathcal{F}_{n,m}\langle I,F \rangle(\cdot):\mathcal{P}^{n,m}\rightarrow \mathcal{P}^{n,m}\langle I,F \rangle$ that applies the same effect to a ratio of polynomials, where
$\mathcal{P}^{n}$ is the
space of polynomials of $n$-th order, $\mathcal{P}^{n}\langle I,F
\rangle$ is the space of polynomials with coefficients in
$\mathbb{R}\langle I,F \rangle$, and $\mathcal{P}^{n,m}$,  $\mathcal{P}^{n,m}\langle I,F
\rangle$ are their rational equivalent.

@@

%------------------------------------
\subsection{Stability under FWL Representation, Quantisation Noise and Parametric Uncertainties}
\label{sec:stability}
%------------------------------------

Sound synthesis of the digital controller requires the consideration of the effect of FWL on the controller, 
and of quantization noises in the closed-loop system.  
Let the quantizer $Q1$ (ADC) be the source of a white noise $\nu_{1}$, 
and $Q2$ (DAC) be the source of a white noise $\nu_{2}$.  The following equation
models the system in Figure~\ref{fig:sampledsystem}, encompassing the
parametric uncertainties $\Delta \vec{G}$ and the FWL effects on
the controller $\hat{C}(z)$:
%
\begin{equation}
\hat{Y}(z)=\nu_{1}(z)+\hat{G}(z)\cdot C(z)\cdot R(z)+\hat{G}(z)\cdot\nu_{2}(z)-\hat{G}(z)\cdot \hat{C}(z)\cdot \hat{Y}(z). 
\end{equation}
%
The above can be rewritten as follows:
%
\begin{equation}
\label{eq:outputfunctions}
\hat{Y}(z)=H_{1}(z)\cdot\nu_{1}(z)+H_{2}(z)\cdot\nu_{2}(z)+H_{3}(z)\cdot R(z),
\end{equation}
%
where
%
\begin{equation*}
H_{1}(z)=\frac{1}{1+\hat{G}(z)\cdot \hat{C}(z)}, 
\end{equation*}
%
\begin{equation*}
H_{2}(z)=\frac{\hat{G}(z)}{1+\hat{G}(z)\cdot \hat{C}(z)}, 
\end{equation*}
%
\begin{equation*}
H_{3}(z)=\frac{\hat{G}(z)\cdot \hat{C}(z)}{1+\hat{G}(z)\cdot \hat{C}(z)}. 
\end{equation*}

\begin{myassumption}
\label{whitenoise}
%
The quantization noises $\nu_{1}$ (from $Q1$) and $\nu_{2}$ (from $Q2$) are
uncorrelated white noises and their amplitudes are always bounded by the
half of quantization step~\cite{astrom1997computer}, i.e., $\vert \nu_{1}
\vert \leq \frac{q_{1}}{2}$ and $\vert \nu_{2} \vert \leq \frac{q_{2}}{2}$,
where $q_{1}$ and $q_{2}$ are the quantization steps of ADC and DAC, respectively.
% 
\end{myassumption}

@@

Next, we describe the specific property that we pass to the program
synthesizer as the specification $\sigma$.  There are two different
algorithms in our verification engine that can be used for stability
~\cite{daes20161, Bessa16}, one based on Schur's decomposition
and another one based on Jury's criterion~\cite{astrom1997computer}.  Here,
we choose Jury's method~\cite{astrom1997computer}, due to its efficiency,
and we use this method to check the stability in the $z$-domain for the
characteristic polynomial $S(z)$ defined in~\eqref{eq:internal_stab_lemma}.
%
% of the matrix $\left( \begin{array}{c} Y \\ U \end{array}\right)$, where $Y(z)$ is the closed-loop system output and U(z) is the controller output. 
%% \red{[unclear: what matrix? We need to clarify the relationship
%% between the char. poly. $S(z)$ and the later quantities $\Delta
%% N_{G}(z), \Delta D_{G}(z)$ and corresponding controller's
%% structure.]}.
%
We~consider the following form for $S(z)$:
%
\begin{equation*}
S(z) = a_0z^N+a_1z^{N-1}+\cdots+a_{N-1}z+a_N=0, a_0\neq0. 
\end{equation*}
%\cdsay{is this for the matrix $\left( \begin{array}{c} Y \\ U \end{array}\right)$?}

@@

The {\sc synthesize} phase uses BMC to 
compute a solution $C(z)$. % An important remark is that this 
% controller is synthesised in the $\mathbb{R}\langle I,F \rangle$
% domain.
There are two approaches for the {\sc verify} phase.
% based on the described model.  
The first approach uses interval arithmetic \cite{moore1966interval}
to represent the coefficients
$[{c}_i-\Delta_p{c}_i-\Delta_b{c}_i,\,\allowbreak
{c}_i+\Delta_p{c}_i-\Delta_b{c}_i+(2^{-F_p})]$ 
and rounds operations outwards during calculations.  This effectively allows
us to simultaneously evaluate the whole collection of plants $\hat{G}(s)$
plus the effects of numeric calculations in the engine.  Synthesized
controllers using this verifier will always be stable for all plants in the
family.
% and only generate counterexamples when the precision of the
%engine is insufficient (\emph{ie} no stable plant could be found). 
Our preliminary experiments showed that a synthesis approach
using this verification engine has poor performance and we
%Given the high computational cost of this
%approach, 
therefore designed a second approach, which we describe below.
Our experimental results described in Section~\ref{sec:experiments}
%try both approaches and provide the time for the fastest. 
show that the speedup yielded by the second approach 
is in most cases of at least two orders of magnitude.
